---
title: "R Notebook"
output: html_notebook
---
```{r}
library(dplyr)
library(tidyverse)
```
# read in the 2 files
```{r}
unemployment = read.csv('../data/UNRATE.csv')
```
```{r}
inflation = read.csv('../data/PCEPILFE.csv')
```

# join the 2 files on the "DATE" column
```{r}
unemploymentInflation = left_join(unemployment, inflation, by = "DATE") 
```

#drop the rows where there is no inflation data
```{r}
unemploymentInflation <- unemploymentInflation[-(1:132),]
```

# rename the UNRATE and PCEPILFE columns; match syntax in DATE column
```{r}
unemploymentInflation = unemploymentInflation %>% 
              transmute(
              date = DATE,
              unemployment_rate = UNRATE,
              core_PCE_index = PCEPILFE)
```

# create a chart graphing unemployment vs inflation over time
```{r}
ggplot(unemploymentInflation, aes(x=unemployment_rate, y=core_PCE_index)) + geom_point()
```
```{r}
library("readxl")
```
# read in the excel files
```{r}
Pre_1970 = read_csv('../data/Pre_1970.csv')
Post_1970 = read_csv('../data/Post_1970.csv')  
```
# concatenate the 2 tables
```{r}
combined <- rbind(Pre_1970, Post_1970)
```
# rename columns
```{r}
unempInfla = combined %>% 
              transmute(
              date = observation_date,
              inflation_rate = PCECTPI_PC1,
              unemployment_rate = UNRATE) 
```
# create a chart graphing unemployment vs inflation over time
```{r}
ggplot(unempInfla, aes(x=unemployment_rate, y=inflation_rate)) + geom_point() 
```
# redo initial code with more complete inflation data
```{r}
unemploymentNew = read.csv('../data/UNRATE.csv')
inflationNew = read.csv('../data/pce_complete.csv') 
```

# rename column in inflationNew
```{r}
inflationNew = inflationNew %>% 
              transmute(
              date = DATE,
              inflation_rate = DPCCRV1Q225SBEA)
```

#combine unemploymentInflation and inflationNew
# graph with new data
```{r}
ggplot(unempInfla, aes(x=unemployment_rate, y=inflation_rate)) + geom_point() 
```



