---
title: "R Notebook"
output: html_notebook
---
```{r}
library(dplyr)
library(tidyverse)
```
# read in files using more complete quarterly data
```{r}
unemploymentNew = read.csv('../data/unrate_quarterly.csv')
inflationNew = read.csv('../data/pce_complete.csv') 
```

# rename columns in unemploymentNew
```{r}
unemploymentNew = unemploymentNew %>% 
              transmute(
              date = DATE,
              unemployment_rate = LRUNTTTTUSQ156S)
```
# rename columns in inflationNew
```{r}
inflationNew = inflationNew %>% 
              transmute(
              date = DATE,
              inflation_rate = DPCCRV1Q225SBEA)
```

#combine unemploymentInflation and inflationNew
```{r}
unemploymentandinflation <- merge(unemploymentNew, inflationNew) 
```

# graph unemployment and inflation using all available data
```{r, message=FALSE}
ggplot(unemploymentandinflation, aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation, 1960 - Present") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)") 
```
# plot by 10-year periods
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1960-01-01" & unemploymentandinflation$date<="1969-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation in the 1960's") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1970-01-01" & unemploymentandinflation$date<="1979-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation in the 1970's") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1980-01-01" & unemploymentandinflation$date<="1989-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation in the 1980's") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1990-01-01" & unemploymentandinflation$date<="1999-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation in the 1990's") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2000-01-01" & unemploymentandinflation$date<="2009-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation in the early 2000's") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2010-01-01" & unemploymentandinflation$date<="2022-07-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation since 2010") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
# plot by 5-year periods
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1960-01-01" & unemploymentandinflation$date<="1964-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1960 - 1964") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1965-01-01" & unemploymentandinflation$date<="1969-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1965 - 1969") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1970-01-01" & unemploymentandinflation$date<="1974-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1970 - 1974") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1975-01-01" & unemploymentandinflation$date<="1979-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1975 - 1979") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1980-01-01" & unemploymentandinflation$date<="1984-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1980 - 1984") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1985-01-01" & unemploymentandinflation$date<="1989-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1985 - 1989") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1990-01-01" & unemploymentandinflation$date<="1994-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1990 - 1994") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1995-01-01" & unemploymentandinflation$date<="1999-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1995 - 1999") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2000-01-01" & unemploymentandinflation$date<="2004-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2000 - 2004") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2005-01-01" & unemploymentandinflation$date<="2009-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2005 - 2009") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2010-01-01" & unemploymentandinflation$date<="2014-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2010 - 2014") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2015-01-01" & unemploymentandinflation$date<="2022-07-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2015 - 2022") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```

# plot by 3-year periods
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1960-01-01" & unemploymentandinflation$date<="1962-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1960-1962") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1963-01-01" & unemploymentandinflation$date<="1965-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1963-1965") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1966-01-01" & unemploymentandinflation$date<="1968-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1966-1968") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1969-01-01" & unemploymentandinflation$date<="1971-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm,se=FALSE  ) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1969-1971") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1972-01-01" & unemploymentandinflation$date<="1974-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1972-1974") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1975-01-01" & unemploymentandinflation$date<="1977-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1975-1977") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1978-01-01" & unemploymentandinflation$date<="1980-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1978-1980") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1981-01-01" & unemploymentandinflation$date<="1983-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1981-1983") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1984-01-01" & unemploymentandinflation$date<="1986-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1984-1986") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1987-01-01" & unemploymentandinflation$date<="1989-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1987-1989") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1990-01-01" & unemploymentandinflation$date<="1992-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1990-1992") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1993-01-01" & unemploymentandinflation$date<="1995-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1993-1995") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1996-01-01" & unemploymentandinflation$date<="1998-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1996-1998") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1999-01-01" & unemploymentandinflation$date<="2001-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 1999-2001") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2002-01-01" & unemploymentandinflation$date<="2004-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2002-2004") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2005-01-01" & unemploymentandinflation$date<="2007-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2005-2007") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2008-01-01" & unemploymentandinflation$date<="2010-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2008-2010") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2011-01-01" & unemploymentandinflation$date<="2013-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2011-2013") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2014-01-01" & unemploymentandinflation$date<="2016-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2014-2016") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2017-01-01" & unemploymentandinflation$date<="2019-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2017-2019") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
```{r, message=FALSE}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2020-01-01" & unemploymentandinflation$date<="2022-10-01"),], aes(x=unemployment_rate, y=inflation_rate)) + geom_point(pch=16, cex=2) + geom_smooth(method=lm, se=FALSE) + geom_point(color = "orange") + theme_bw() + ggtitle("Unemployment vs. Inflation 2020-2022") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Unemployment Rate (%)") + ylab("Inflation Rate (PCEPI, %)")
```
## graph unemployment and inflation on the same axis over same time periods as scatter plots

```{r}
ggplot(unemploymentandinflation, aes(x = lubridate::as_date(date))) + 
  geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1960 - Present") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1960-01-01" & unemploymentandinflation$date<="1969-10-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1960 - 1969") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")  
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1970-01-01" & unemploymentandinflation$date<="1979-10-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1970 - 1979") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1980-01-01" & unemploymentandinflation$date<="1989-10-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1980 - 1989") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")  
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1990-01-01" & unemploymentandinflation$date<="1999-10-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1990 - 1999") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2000-01-01" & unemploymentandinflation$date<="2009-10-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2000 - 2009") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2010-01-01" & unemploymentandinflation$date<="2022-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2010 - Present") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1960-01-01" & unemploymentandinflation$date<="1964-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1960 - 1964") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1965-01-01" & unemploymentandinflation$date<="1969-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1965 - 1969") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1970-01-01" & unemploymentandinflation$date<="1974-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1970 - 1974") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1975-01-01" & unemploymentandinflation$date<="1979-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1975 - 1979") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1980-01-01" & unemploymentandinflation$date<="1984-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1980 - 1984") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1985-01-01" & unemploymentandinflation$date<="1989-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1985 - 1989") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1990-01-01" & unemploymentandinflation$date<="1994-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1990 - 1994") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1995-01-01" & unemploymentandinflation$date<="1999-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1995 - 1999") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2000-01-01" & unemploymentandinflation$date<="2004-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2000 - 2004") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2005-01-01" & unemploymentandinflation$date<="2009-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2005 - 2009") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2010-01-01" & unemploymentandinflation$date<="2014-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2010 - 2014") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2015-01-01" & unemploymentandinflation$date<="2022-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2015 - Present") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1960-01-01" & unemploymentandinflation$date<="1962-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1960 - 1962") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1963-01-01" & unemploymentandinflation$date<="1965-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1963 - 1965") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1966-01-01" & unemploymentandinflation$date<="1968-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1966 - 1968") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1969-01-01" & unemploymentandinflation$date<="1971-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1969 - 1971") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1972-01-01" & unemploymentandinflation$date<="1974-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1972 - 1974") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1975-01-01" & unemploymentandinflation$date<="1977-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1975 - 1977") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1978-01-01" & unemploymentandinflation$date<="1980-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1978 - 1980") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1981-01-01" & unemploymentandinflation$date<="1983-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1981 - 1983") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1984-01-01" & unemploymentandinflation$date<="1986-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1984 - 1986") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1987-01-01" & unemploymentandinflation$date<="1989-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1987 - 1989") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1990-01-01" & unemploymentandinflation$date<="1992-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1990 - 1992") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1993-01-01" & unemploymentandinflation$date<="1995-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1993 - 1995") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1996-01-01" & unemploymentandinflation$date<="1998-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1996 - 1998") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="1999-01-01" & unemploymentandinflation$date<="2001-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 1999 - 2001") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2002-01-01" & unemploymentandinflation$date<="2004-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2002 - 2004") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2005-01-01" & unemploymentandinflation$date<="2007-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2005 - 2007") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2008-01-01" & unemploymentandinflation$date<="2010-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2008 - 2010") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2011-01-01" & unemploymentandinflation$date<="2013-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2011 - 2013") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2014-01-01" & unemploymentandinflation$date<="2016-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2014 - 2016") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %") 
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2017-01-01" & unemploymentandinflation$date<="2019-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2017 - 2019") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```
```{r}
ggplot(unemploymentandinflation[which(unemploymentandinflation$date>="2020-01-01" & unemploymentandinflation$date<="2022-07-01"),], aes(x = lubridate::as_date(date))) + geom_line(aes(y = unemployment_rate), color = "blue", linewidth=0.8) + geom_line(aes(y = inflation_rate), color = "orange", linewidth=0.8) + theme_bw() + ggtitle("Unemployment (blue) vs. Inflation (orange) 2020 - 2022") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Date") + ylab("Unemployment %,  Inflation (PCEPI) %")
```








